---
title: "Machine Learning Competition: Report on Forest Cover Type Prediction"
author: "Fernandez, María, McIver, Jordan, Besanson, Gaston"
output: rmarkdown::tufte_handout
---

# Introduction

\newthought{This report} has the purpose to explain our experience on the Forest Cover Type Prediction's Competition.^[https://inclass.kaggle.com/c/prediction-of-a-forest-covertype] This experience starts with `Data exploration and Feature Creation`, `Pre-Processing`, `Machine Learning Methods`, `Results over this dataset`, `Conclusion`.

# Data Description^[The Feature creation is explained in the attached powerpoint presentation.]

how many features. how many classes. Table explaning what is the class more populated. In test and trainning set.
IDEAS?

# Pre-Processing

We try PCA. Did not work. Why? Explain the pre-processing. Why we did it.

# Machine Learning Methods

## Random Forest
What package we used. Why we select this method. Literature

## Support Vector Machine
What package we used. Why we select this method. Literature

## Gradient Boosting Methods
What package we used. Why we select this method. Literature

## Deep Learning. Just for fun
What package we used. Why we select this method. Literature

# Results

Decide to do Cross Validation. The accuracy in the training. accuarcy in the leaderboard. ROC. Tables
IDEAS?

If we have the 3 methods.. we can do a voting......

# Conclusion


----------------------------------------------------------------

\begin{marginfigure}
$$\frac{d}{dx}\left( \int_{0}^{x} f(u)\,du\right)=f(x).$$
\caption{An equation}
\end{marginfigure}

Note the use of the `\caption` command to add additional text below the equation.

## Full Width Figures

You can arrange for figures to span across the entire page by using the `fig.fullwidth` chunk option. 

```{r, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "Full width figure"}
library(ggplot2)
qplot(wt, mpg, data=mtcars, colour=factor(cyl))
```


# Tables

You can use the **xtable** package to format \LaTeX\ tables that integrate well with the rest of the Tufte handout style. Note that it's important to set the `xtable.comment` and `xtable.booktabs` options as shown below to ensure the table is formatted correctly for inclusion in the document.

```{r, results='asis'}
library(xtable)
options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)
xtable(head(mtcars[,1:6]), caption = "First rows of mtcars")
```